### 2020-04-26 serial.tex
spell-checking

We provide a simple, heuristic way of assessing potential biases in the estimate of the mean initial forward serial interval and therefore $\mathcal R$ retrospectively.
Once serial intervals have been observed after the epidemic has been sufficiently progressed, we can group observed serial intervals by their primary cohort times (i.e., symptom onset times of infectors).
Then, we can compare how estimates of the mean serial interval as well as $\mathcal R$ change as we include more recent cohorts into the analysis (see `cohort-averaged' in \fref{obsrR}).
During the exponential growth phase, the estimates of the mean serial interval and $\mathcal R$ are consistent with the target value;
adding more data allows us to make more precise inference during this period.
However, the cohort-averaged estimates decreases rapidly soon after the exponential growth period.
This approach allows us to characterize potential biases in the estimates of $\mathcal R$ cuased by changes in the forward serial-interval distributions.

@@

Desipte these limitations, our analysis of serial intervals of COVID-19 provide further support for our theoretical framework, demonstrating temporal variation in serial intervals and their effects on the estimates of $\mathcal R$.
To our knowledge, most, if not all, existing estimates of the serial-intervals of COVID-19 implicitly or explicitly assume that the serial-interval distributions remain constant throughout the course of an epidemic \citep{du2020serial, he2020temporal, nishiura2020serial,tindale2020transmission,zhao2020estimating,zhang2020evolving}.
Our study provides a rationale for reassessing estimates of serial-interval distributions of the COVID-19 pandemic.

@@

We simulate the renewal equation model using a discrete-time approximation:
\begin{equation}
\begin{aligned}
i(t) &= \mathcal R_0 S(t-\Delta t) \sum_{m=1}^{\ell} i(t-\Delta t m) \hat{g}(\Delta t m) \\
S(t) &= S(t-\Delta t) - i(t)
\end{aligned}
\end{equation}
where $\hat{g}$ is a discrete-time intrinsic generation-interval distribution that satisfies the following:
\begin{equation}
\hat{g}(\Delta t m) = \frac{g(\Delta t m)}{\sum_{i=1}^\ell g(\Delta t m)}, \quad m=1, \dots, \ell
\end{equation}
The continuous-time intrinsic generation-interval distribution is parameterized using a log-normal distribution (Table 1). We define the forward incubation period distribution in a similar manner:
\begin{equation}
\hat{k}(\Delta t m) = \frac{k(\Delta t m)}{\sum_{i=1}^\ell k(\Delta t m)}, \quad m=1, \dots, \ell,
\end{equation}
where its continuous-time analog is also based on a log-normal distribution.
For brevity, we assume that the forward incubation periods and intrisic generation intervals are independent:
\begin{equation}
\hat{h}(\Delta t m, \Delta t n) = \hat{k}(\Delta t m)\hat{g}(\Delta t n), \quad m,n=1, \dots, \ell.
\end{equation}
We use $\Delta t = 0.025\,\textrm{days}$ and $\ell=2001$ for discretization steps.

@@

We run stochastic simulations of the renewal equation model using an individual-based model on a fully connected network (i.e., homogeneous population) based on the Gillespie algorithm that we devloped earlier \citep{park2019inferring}.
First, we initialize an epidemic with $I(0)$ infected individuals (nodes) in a fully connected network of size $N$. 
For each initially infected individual, we draw number of infectious contacts from a Poisson distribution with the mean of $\mathcal R_0$ and the corresponding generation intervals for each contact from a log-normal distribution (Table 1).
Contactees are uniformly sampled from the total population.
All contactees are sorted into event queues based on their infection time.
We update the current time to the infection time of the first person in the queue.
Then, the first person in the queue makes contacts based on the Poisson offspring distribution described earlier and their contactees are added to the sorted queue.
Whenever contactees are aded to the sorted queue, we remove all duplicated contacts (but keep the first one) as well as contacts made to individuals that have already been infected.
Simulations continue until there are no more individuals in the queue.
We simulate 10 epidemics with $I(0)=10$ and $N$=40,000.

@@

We are left to show that 
\begin{equation}
\int_0^{\infty} \exp(-r\tau) g(\tau) \mathrm{d}\tau = \int_{-\infty}^\infty \exp(-r\tau) \int_{\max{(0,-\alpha_2)}}^{\infty} \exp(-r\alpha_1) h(\alpha_1, \alpha_2 + \alpha_1) \mathrm{d}\alpha_1 \mathrm{d}\tau,
\end{equation}
where the intrisic generation-interval distribution $g$ is also a marginal probability distribution of $f$:
\begin{equation}
g(\tau) = \int_0^\infty h(\alpha_1, \tau)  \mathrm{d} \alpha_1.
\end{equation}
Let $\sigma = \alpha_1 + \tau$. Then, by change of variables, it immediately follows that
\begin{equation}
\begin{aligned}
&\int_{-\infty}^{\infty} \exp(-r\tau) \int_{\max(0, -\tau)}^\infty \exp(- r \alpha_1) h(\alpha_1, \tau+\alpha_1) \mathrm{d} \alpha_1\, \mathrm{d}\tau\\
&=\int_{0}^{\infty} \int_{0}^\infty \exp(- r \sigma) h(\alpha_1, \sigma) \mathrm{d} \alpha_1\, \mathrm{d}\sigma\\
&=\int_{0}^{\infty} \exp(-r\tau) g(\tau) \mathrm{d}\tau
\end{aligned}
\end{equation}
Therefore, the initial forward serial--interval distribution and the intrinsic generation-interval distribution give the same link between $r$ and $\mathcal R$.

>>>>

We provide a simple, heuristic way of assessing potential biases in the estimate of the mean initial forward serial interval and therefore $\mathcal R$ retrospectively.
Once serial intervals have been observed after the epidemic has been sufficiently progressed, we can group observed serial intervals by their primary cohort times (i.e., symptom onset times of infectors).
Then, we can compare how estimates of the mean serial interval as well as $\mathcal R$ change as we include more recent cohorts into the analysis (see `cohort-averaged' in \fref{obsrR}).
During the exponential growth phase, the estimates of the mean serial interval and $\mathcal R$ are consistent with the target value;
adding more data allows us to make more precise inference during this period.
However, the cohort-averaged estimates decreases rapidly soon after the exponential growth period.
This approach allows us to characterize potential biases in the estimates of $\mathcal R$ caused by changes in the forward serial-interval distributions.

@@

Despite these limitations, our analysis of serial intervals of COVID-19 provide further support for our theoretical framework, demonstrating temporal variation in serial intervals and their effects on the estimates of $\mathcal R$.
To our knowledge, most, if not all, existing estimates of the serial-intervals of COVID-19 implicitly or explicitly assume that the serial-interval distributions remain constant throughout the course of an epidemic \citep{du2020serial, he2020temporal, nishiura2020serial,tindale2020transmission,zhao2020estimating,zhang2020evolving}.
Our study provides a rationale for reassessing estimates of serial-interval distributions of the COVID-19 pandemic.

@@

We simulate the renewal equation model using a discrete-time approximation:
\begin{equation}
\begin{aligned}
i(t) &= \mathcal R_0 S(t-\Delta t) \sum_{m=1}^{\ell} i(t-\Delta t m) \hat{g}(\Delta t m) \\
S(t) &= S(t-\Delta t) - i(t)
\end{aligned}
\end{equation}
where $\hat{g}$ is a discrete-time intrinsic generation-interval distribution that satisfies the following:
\begin{equation}
\hat{g}(\Delta t m) = \frac{g(\Delta t m)}{\sum_{i=1}^\ell g(\Delta t m)}, \quad m=1, \dots, \ell
\end{equation}
The continuous-time intrinsic generation-interval distribution is parameterized using a log-normal distribution (Table 1). We define the forward incubation period distribution in a similar manner:
\begin{equation}
\hat{k}(\Delta t m) = \frac{k(\Delta t m)}{\sum_{i=1}^\ell k(\Delta t m)}, \quad m=1, \dots, \ell,
\end{equation}
where its continuous-time analog is also based on a log-normal distribution.
For brevity, we assume that the forward incubation periods and intrinsic generation intervals are independent:
\begin{equation}
\hat{h}(\Delta t m, \Delta t n) = \hat{k}(\Delta t m)\hat{g}(\Delta t n), \quad m,n=1, \dots, \ell.
\end{equation}
We use $\Delta t = 0.025\,\textrm{days}$ and $\ell=2001$ for discretization steps.

@@

We run stochastic simulations of the renewal equation model using an individual-based model on a fully connected network (i.e., homogeneous population) based on the Gillespie algorithm that we developed earlier \citep{park2019inferring}.
First, we initialize an epidemic with $I(0)$ infected individuals (nodes) in a fully connected network of size $N$. 
For each initially infected individual, we draw number of infectious contacts from a Poisson distribution with the mean of $\mathcal R_0$ and the corresponding generation intervals for each contact from a log-normal distribution (Table 1).
Contactees are uniformly sampled from the total population.
All contactees are sorted into event queues based on their infection time.
We update the current time to the infection time of the first person in the queue.
Then, the first person in the queue makes contacts based on the Poisson offspring distribution described earlier and their contactees are added to the sorted queue.
Whenever contactees are added to the sorted queue, we remove all duplicated contacts (but keep the first one) as well as contacts made to individuals that have already been infected.
Simulations continue until there are no more individuals in the queue.
We simulate 10 epidemics with $I(0)=10$ and $N$=40,000.

@@

We are left to show that 
\begin{equation}
\int_0^{\infty} \exp(-r\tau) g(\tau) \mathrm{d}\tau = \int_{-\infty}^\infty \exp(-r\tau) \int_{\max{(0,-\alpha_2)}}^{\infty} \exp(-r\alpha_1) h(\alpha_1, \alpha_2 + \alpha_1) \mathrm{d}\alpha_1 \mathrm{d}\tau,
\end{equation}
where the intrinsic generation-interval distribution $g$ is also a marginal probability distribution of $f$:
\begin{equation}
g(\tau) = \int_0^\infty h(\alpha_1, \tau)  \mathrm{d} \alpha_1.
\end{equation}
Let $\sigma = \alpha_1 + \tau$. Then, by change of variables, it immediately follows that
\begin{equation}
\begin{aligned}
&\int_{-\infty}^{\infty} \exp(-r\tau) \int_{\max(0, -\tau)}^\infty \exp(- r \alpha_1) h(\alpha_1, \tau+\alpha_1) \mathrm{d} \alpha_1\, \mathrm{d}\tau\\
&=\int_{0}^{\infty} \int_{0}^\infty \exp(- r \sigma) h(\alpha_1, \sigma) \mathrm{d} \alpha_1\, \mathrm{d}\sigma\\
&=\int_{0}^{\infty} \exp(-r\tau) g(\tau) \mathrm{d}\tau
\end{aligned}
\end{equation}
Therefore, the initial forward serial--interval distribution and the intrinsic generation-interval distribution give the same link between $r$ and $\mathcal R$.
